---
title: "Johnny Rodriguez Project 2"
author: "Johnny Rodriguez"
date: "2022-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Prepare Environment
Loads the libraries

```{r}
library (dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library (ggplot2)
```

## NY Times Bestsellers for Manga Books
This data looks at Manga books from the New York Times Best Seller list from 2010-2017.  The data is transformed through a series of step for further analysis.


```{r}

# Imports the CSV of all bestsellters
nytbestsellers <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/nytbestsellers.csv", header = TRUE, sep = ',', na.strings="", fill = TRUE)
knitr::kable(head(nytbestsellers))

# Filters the initial large best sellers list for manga books and a subset of columns to make the data more manageable
nytmanga <- nytbestsellers %>% 
  filter(list_name_encoded == "manga") %>% 
  select(published_date, list_name_encoded, rank, title, author, price, weeks_on_list )


# Converts the character date columns into a real date column; adds a separate column that permits a yearly plot by using the floor date function.
nytmanga <- nytmanga %>% 
  mutate(published_date = ymd(published_date),
         pubfloordate = floor_date(published_date, unit = "year" ))


# Converts the book title column into separate columns to split between the main title and the volume number.
nytmanga <- nytmanga %>%
  mutate(main_title = ifelse(str_detect(title, 'VOL.'),
                      str_match(title, '(.*),\\sVOL.')[,2],
                      title),
         VolumeNo = str_match(title, 'VOL.\\s{1,}(\\d+)')[,2])

# Splits the author column into a first and second author if one was noted.
nytmanga <- nytmanga %>%
  separate(col = author, sep = ' and ', into = c('first_author', 'second_author'), remove = FALSE)

knitr::kable(head(nytmanga))
```

#### Conclusion - Manga Best Sellers
Between 2010-2017, the  majority of Mango books spent about 1.25 weeks on the NYT Best Seller list.  The number of manga books on the best seller list for at least 1 week peaked in 2012.


```{r}

ggplot(nytmanga, aes(x=weeks_on_list)) + geom_bar() +
  ggtitle("Manga NYT Best Sellers Number of Weeks on List By Year 2010 - 2017")+
  scale_x_continuous(limits = c(0,10)) +
  facet_wrap(~pubfloordate)

```

## Global Happiness Scores 2015-2022
This data looks at the global happiness score between 2015 - 2022.  This information is available on a number of csv files that required joining and transformation.



```{r}
# Imports the Happiness data for each year
happy2015 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2015.csv", na.strings=c("","NA"))
happy2016 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2016.csv", na.strings=c("","NA"))
happy2017 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2017.csv", na.strings=c("","NA"))
happy2018 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2018.csv", na.strings=c("","NA"))
happy2019 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2019.csv", na.strings=c("","NA"))
happy2020 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2020.csv", na.strings=c("","NA"))
happy2021 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2021.csv", na.strings=c("","NA"))
happy2022 <- read.csv("https://raw.githubusercontent.com/johnnydrodriguez/data607_project2/main/2022.csv", na.strings=c("","NA"))

# Example yearly happiness data
knitr::kable(head(happy2015))

# Selects the happiness score from 2015 to create global happiness data frame; the data will only look at the happiness score  per country.
globalhappy <- happy2015 %>% 
  select(Country, Happiness.Score) %>%
  rename(Happiness.Score.2015 = Happiness.Score) %>% 
  
# Joins the 2016 happiness score.
  full_join(happy2016, by = 'Country') %>% 
  select(Country, Happiness.Score.2015, Happiness.Score) %>%
  rename(Happiness.Score.2016 = Happiness.Score) %>% 
  
# Joins the 2017 happiness score.  
  full_join(happy2017, by = 'Country') %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score) %>%
  rename(Happiness.Score.2017 = Happiness.Score) %>% 
  
# In 2018, the column names change.  Joins based on new country and new score column names.  
  full_join(happy2018, by = c("Country" = "Country.or.region")) %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score.2017, Score) %>%
  rename(Happiness.Score.2018 = Score) %>%
  
# Joins 2019 data keeping the 2018 column names format.
  full_join(happy2019, by = c("Country" = "Country.or.region")) %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score.2017,Happiness.Score.2018, Score) %>%
  rename(Happiness.Score.2019 = Score) %>%
  
# In 2020, the column names for country and score change again.  Joins the 2020 happiness score.  
  full_join(happy2020, by = c("Country" = "Country.name")) %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score.2017,Happiness.Score.2018, Happiness.Score.2019, Ladder.score) %>%
  rename(Happiness.Score.2020 = Ladder.score) %>% 

# Joins 2021 happiness dta keeping the 2020 column name format    
  full_join(happy2021, by = c("Country" = "Country.name")) %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score.2017,Happiness.Score.2018, Happiness.Score.2019, Happiness.Score.2020, Ladder.score) %>%
  rename(Happiness.Score.2021 = Ladder.score) %>% 
  
# In 2022, the column names for country and score change again.  Joins the 2022 happiness score. 
  full_join(happy2022, by = 'Country') %>% 
  select(Country, Happiness.Score.2015, Happiness.Score.2016,Happiness.Score.2017,Happiness.Score.2018, Happiness.Score.2019, Happiness.Score.2020, Happiness.Score.2021, Happiness.score) %>%
  rename(Happiness.Score.2022 = Happiness.score)

# In 2022, the Happiness score format is changed to include a comma instead of a decimals.  The values are converted to numeric and then changed to the same scale as the previous years.
globalhappy$Happiness.Score.2022 <- as.numeric(gsub(",","", globalhappy$Happiness.Score.2022))
globalhappy <- globalhappy %>% 
  mutate(Happiness.Score.2022 = Happiness.Score.2022 / 1000)

# The average happines score for each country is calculated and NAN value is removed.
globalhappy <- globalhappy %>% 
  mutate(AvgHappiness = (rowMeans(select(globalhappy, "Happiness.Score.2015", "Happiness.Score.2016","Happiness.Score.2017","Happiness.Score.2018", "Happiness.Score.2019", "Happiness.Score.2020", "Happiness.Score.2021","Happiness.Score.2022"), na.rm = TRUE))) %>% 
  filter(!is.na(AvgHappiness))

knitr::kable(head(globalhappy))

```

#### Conclusion - Distribution of Global Happiness scores

Average global happiness scores between 2015- 2022 appear to be normally distributed.  To confirm a density plot which places the data below the curve and a QQ distribution line where all of the points appear to fall near the line.

```{r}
# Mean and Standard Deviation for Average Global Happiness scores
happymean <- mean(globalhappy$AvgHappiness)
happysd   <- sd(globalhappy$AvgHappiness)
happymean
happysd

#Density Distribtion Plot for Average Happiness score
ggplot(globalhappy, aes(x = AvgHappiness)) +geom_blank() +geom_histogram(aes(y = ..density..)) +stat_function(fun = dnorm, args = c(mean = happymean, sd = happysd), col = "tomato")

## QQ Distibrution line to evaluate normal distribution of happiness scores
ggplot(globalhappy, aes(sample = AvgHappiness)) + geom_line(stat = "qq")+ ggtitle("Global Average Happiness QQ Distribution Line")

```


